Beyond Adaptive Web Design

### 2014-02-20 paper/paper.tex
refactor(paper): refactor paper parts

The basic approach how to provide an optimal viewing experience is to use ``Responsive web design'' which comes with fluid grids, flexible images and CSS3 media queries~\cite{responsive}. While creating flexible layouts is important, there's a lot more that requires exceptional adaptive web experiences. The principles of Adaptive Design are: ubiquity, flexibility, performance, enhancement and future-friendliness~\cite{adaptivesxsw}.

@@

Web browsers on TVs are actually really good but the input is terrible so no one uses them. Input is much more important to interface design than screen size. The input defines what a design needs to do in order to accomplish a task~\cite{adaptiveinput}.

We also have wearable computers. The current Google Glass interface is limited to a floating rectangle above head. We don't have full range of motion and, as a result, no spatial cues or depth in the interface.

@@

\paragraph{Touch} % (fold)
\label{sub:touch}

@@

% paragraph touch (end)

\paragraph{Speech} % (fold)
\label{sub:speech}

@@

% paragraph speech (end)

\paragraph{Device motion} % (fold)
\label{sub:device_motion}

@@

% paragraph device_motion (end)

\paragraph{Video Motion} % (fold)
\label{sub:video_motion}

@@

% paragraph video_motion (end)

@@

We all should care about page speed and mobile performance because people need fast experiences on the Web. There's lot of studies that highlight the importance of speed. More than half of people with a bad loading experience on mobile, won't come back~\cite{performancebrowsernetworking}. 73\% of mobile internet users say they've encountered Web pages that are too slow. A 1 second delay can result in a 7\% reduction in conversions~\cite{pagespeed}. Speed is an esencial part of our strategy and we should care about it. We can quantify its impact on the metrics and track it.

There are many different types of internet connections. Mobile network speeds have gone up but this doesn't help much as page load times tend to plateau as bandwidth increases. Latency is a much bigger contribution to download times. LTE reduces tower latency by several milliseconds but we still need to make a number of connections to get data to our phone~\cite{performancebrowsernetworking}.

So what we should do about it? We can create reuseable components and conditionaly load resources. Web components allow us to use custom HTML elements in the browser. We can wrap up a simple element or an entire application logic within an HTML element~\cite{webcomponents}. We can create Web applications in a reusable, modular and encapuslated way.

@@

We have detected several web componenets and input methods and tried to verify which of them is more useful for the web developers. There have been made series of experiments on both adaptive input methods and web components. To prove the usefulness of the proposed concept we have designed and implemented several reusable modules and components and we have used them in propotype web projects. All of them have been released as open source and are available on GitHub~\footnotemark.

@@

\paragraph{Voice commands} % (fold)
\label{ssub:voice_commands}

@@

% paragraph voice_commands (end)

\paragraph{Gyroscope rotation} % (fold)
\label{ssub:gyroscope_rotation}

@@

\begin{figure}[ht]
    \begin{center}
        \includegraphics[width=0.2\textwidth]{../images/gyroscope}
        \caption{Scrolling in the webpage by gyroscope}
        \label{fig:gyroscope}
    \end{center}
\end{figure}

@@

% paragraph gyroscope_rotation (end)

\paragraph{Video camera motion} % (fold)
\label{ssub:video_camera_motion}

@@

\begin{figure}[ht]
    \begin{center}
        \includegraphics[width=0.8\textwidth]{../images/videomotion}
        \caption{Converting video stream into motion vector}
        \label{fig:videomotion}
    \end{center}
\end{figure}

@@

% paragraph video_camera_motion (end)

@@

There are a lot of commonly used web services that produce unnecessary traffic. Elements like videos or maps are a common part of the most web sites. All of them produces many requests and the web traffic even the website user doesn't want to watch the video or browse the map. There are also native mobile application for that services with better perfomance. So what can we to do? This is a great opportunity to utilize conditional loading to serve the best experience for the right context.

@@

\begin{figure}[ht]
    \begin{center}
        \includegraphics[width=0.8\textwidth]{../images/maps.png}
        \caption{Maps component}
        \label{fig:maps}
    \end{center}
\end{figure}

@@

This approach works quite well for simple use cases. We can show multiple map types, markers or videos. However, more interactive elements require additional consideration. Even then use of an embedded elements directly still might not make sense because of unnecessary traffic, while adaptive techniques reduces it. We can still replace basic static image with richer elements. Adaptive web componets are a decent baseline.

@@

We have received a lot feedback and have usage and popularity results of adaptive input methods. As we have expected the most popular input method is speech input as we can control an entire web application using only voice commands. Gyroscope and video motion are also useful input methods, but there are limited occasions where to use them and can be used as supplementary input method only. All of them however provide a great opportunity to enhance user experience.

Adaptive web components are a great to save an unwonted web traffic. They also icrease user experience because the web page produces less requests and loads faster. However the limitations also exist because many websites need custom elements and design. As a result the adaptive web components produce less interest than adaptive input methods.

More changes are coming, new devices, web APIs and beyond. What will it look like in a couple of years? The devices will be more diversified and have the web access. So we have to prepare the Web to this evolution and provide the best user experience.

>>>>

The basic approach how to provide an optimal viewing experience is to use ``Responsive web design'' which comes with fluid grids, flexible images and CSS3 media queries~\cite{responsive}. While creating flexible layouts is important, there's a lot more that requires remarkable adaptive web experiences. The principles of Adaptive Design are: ubiquity, flexibility, performance, enhancement and future-friendliness~\cite{adaptivesxsw}.

@@

Web browsers on TVs are actually really good but the input is terrible so no one uses them. Input is much more important to interface design than screen size. The input defines what a design needs to do in order to accomplish a task~\cite{adaptiveinput}. We also have wearable computers. The current Google Glass interface is limited to a floating display above head. We don't have full range of motion and, as a result, no spatial cues or depth in the interface.

@@

\subsubsection*{Touch} % (fold)
\label{sub:touch}

@@

% subsubsection touch (end)

\subsubsection*{Speech} % (fold)
\label{sub:speech}

@@

% subsubsection speech (end)

\subsubsection*{Device motion} % (fold)
\label{sub:device_motion}

@@

% subsubsection device_motion (end)

\subsubsection*{Video Motion} % (fold)
\label{sub:video_motion}

@@

% subsubsection video_motion (end)

@@

We all should be interested in page speed and mobile performance because people need fast experiences on the Web. There's lot of studies that highlight the importance of speed. More than half of people with a bad loading experience on mobile, won't come back~\cite{performancebrowsernetworking}. 73\% of mobile internet users say they've encountered Web pages that are too slow, a 1 second delay can result in a 7\% reduction in conversions~\cite{pagespeed}. Speed is an esencial part of our strategy and we should care about it. We can quantify its impact on the metrics and track it.

There are many different types of internet connections. Mobile network speeds have rose but this doesn't help much as page load times are still high as bandwidth increases. Latency is a much bigger fragment to download times. LTE reduces tower latency by several milliseconds but we still need to make a number of connections to download data to our devices~\cite{performancebrowsernetworking}.

So what we should do about it? We can create reuseable components and conditionaly load resources. Web components allow us to use custom HTML elements in the browser. We can enclose a simple element or an entire application logic within an HTML element~\cite{webcomponents}. We can build Web applications in a reusable, commutable and encapuslated way.

@@

We have detected several web componenets and input methods and tried to verify which of them provide better experience for the website users and are also useful for the web developers. There have been made series of experiments on both adaptive input methods and adaptive web components. To prove the usefulness of the proposed concept we have designed and implemented several reusable modules and components and we have used them in propotype web projects. All of them have been released as open source and are available on GitHub~\footnotemark.

@@

\subsubsection*{Voice commands} % (fold)
\label{ssub:voice_commands}

@@

% subsubsection voice_commands (end)

\subsubsection*{Gyroscope rotation} % (fold)
\label{ssub:gyroscope_rotation}

@@

\begin{figure}[ht]
    \begin{center}
        \includegraphics[width=0.15\textwidth]{../images/gyroscope}
        \caption{Scrolling in the webpage by gyroscope}
        \label{fig:gyroscope}
    \end{center}
\end{figure}

@@

% subsubsection gyroscope_rotation (end)

\subsubsection{Video camera motion} % (fold)
\label{ssub:video_camera_motion}

@@

\begin{figure}[ht]
    \begin{center}
        \includegraphics[width=0.5\textwidth]{../images/videomotion}
        \caption{Converting video stream into motion vector}
        \label{fig:videomotion}
    \end{center}
\end{figure}

@@

% subsubsection video_camera_motion (end)

@@

There are a lot of commonly used web services that produce unnecessary traffic. Elements like videos or maps are a common part of the most web sites. All of them produces many requests and the web traffic even the website user doesn't want to watch the video or browse the map. There are also native mobile application for that services with better perfomance which provide full user experience. So what can we to do? This is a great opportunity to utilize conditional loading to serve the best experience for the right context.

@@

\begin{figure}[ht]
    \begin{center}
        \includegraphics[width=0.55\textwidth]{../images/maps.png}
        \caption{Maps component}
        \label{fig:maps}
    \end{center}
\end{figure}

@@

This approach works quite well for simple use cases. We can show multiple map types, markers or videos. However, more interactive elements require additional consideration. Even then use of an embedded elements directly still might not make sense because of unnecessary traffic, while adaptive techniques reduces it. We can still replace basic static image with richer elements.

@@

We have received a lot feedback and have usage and popularity results of adaptive input methods. As we have expected, the most popular input method is speech input since we can control an entire web application using only voice commands. Gyroscope and video motion are also useful input methods, but there are limited occasions where to use them and can be used as supplementary input method only. All of them however provide a great opportunity to enhance user experience.

Adaptive web components are a great to save an unwonted web traffic. They also icrease user experience because the web page produces less requests and loads faster. However the limitations also exist because many websites need custom elements and design. As a result the adaptive web components produce less interest than adaptive input methods, but are the necessary part of the Web.

More changes are coming, new devices, web APIs and beyond. What will it look like in a couple of years? The devices will be more diversified and have the web access. So we have to prepare the Web to this evolution and provide the best user experience, design for many inputs and save the inessential web traffic.
